Standalone Tactics using OpenTheory

### 2012-02-09 paper.tex
fleshing out sec 2.2

Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules.
In HOL4 and HOL Light, we met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
Then we express each primitive inference as a derived rule above the \OpenTheory kernel.
In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
<could be difficult in system with significantly different kernel, eg isabelle>

<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>

<Implementation details to put on the web: simple cgi>

@@

System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
<Comparison to THF?>
Evidential tool bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb}.
Omega.
Natural extension of the integration of automatic tools into interactive systems.

@@

Highly customised tactics for a particular proof system might suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.
By the same token, the article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.

>>>>

Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules.
In HOL4 and HOL Light, we met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
Then we express each primitive inference as a derived rule above the \OpenTheory kernel.
In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
Although Isabelle/HOL uses the LCF architecture, its logical kernel is quite different from \OpenTheory; we therefore expect translating Isabelle/HOL proofs to be more difficult than HOL4 proofs.

To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the ``export'' facility of the PolyML compiler, which creates a standalone executable that runs an ML function.
(For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.)
The situation for HOL Light is more complicated because OCaml does not provide such an ``export'' facility, and HOL Light does not use stored theories.
We used a general purpose checkpointing facility to capture a HOL Light session at an appropriate point (ready to read an article and run a tactic).

@@

System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
<Comparison to THF?>
Evidential tool bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb}.
Omega.
MathServe?
Natural extension of the integration of automatic tools into interactive systems.

@@

Using \OpenTheory for communication, we can

Highly customised tactics for a particular proof system might suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.
<The article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.>
<Tactics where proof search dominates proof replay offer the best prospects, since the search can be done elsewhere (in parallel), the overhead is mainly in transporting and replaying proofs, and queries might be cached>
